---
format: gfm
execute:
  echo: false
  warning: false
---

```{r}
#| include: false
library(tidyverse)
library(glue)
```

# Machine Learning foR Psychologists

[![](https://img.shields.io/badge/Open%20Educational%20Resources-Compatable-brightgreen)](https://creativecommons.org/about/program-areas/education-oer/)
[![](https://img.shields.io/badge/CC-BY--NC%204.0-lightgray)](http://creativecommons.org/licenses/by-nc/4.0/)
[![](https://img.shields.io/badge/Language-R-blue)](http://cran.r-project.org/)

<sub>*Last updated `r Sys.Date()`.*</sub>

This Github repo contains all lesson files for *Machine Learning in R*.
The goal is to impart students with the basic tools to construct, evaluate and compare various **machine learning models, using [`tidymodels`](https://www.tidymodels.org/)**,
based on [_An Introduction to Statistical Learning: with applications in R_](https://www.statlearning.com/).

These topics were taught in the graduate-level course ***Machine Learning for Psychologists*** (Psych Dep., Ben-Gurion University of the Negev; Psych Dep., Tel-Aviv University). This course assumes basic competence in R (importing, regression modeling, plotting, etc.), along the lines of [*Practical Applications in R for Psychologists*](https://github.com/mattansb/Practical-Applications-in-R-for-Psychologists).


**Notes:**

- This repo contains only materials relating to *Practical Applications in R*, and does not contain any theoretical or introductory materials.
- Please note that some code does not work *on purpose*, to force students to learn to debug.

## Setup

You will need:

1. A fresh installation of [**`R`**](https://cran.r-project.org/) (preferably version 4.5.0 or above).
2. [RStudio IDE](https://www.rstudio.com/products/rstudio/download/) or [Positron](https://positron.posit.co/download) (optional, but recommended).
3. The following packages, listed by lesson:


```{r}
script_deps <- function(fl, comment = FALSE) {
  txt <- read.delim(fl, header = FALSE)[[1]] |>
    stringr::str_trim()

  if (isFALSE(comment)) {
    # remove comments from the code
    txt <- stringr::str_remove_all(txt, pattern = "(?<=#).*")
  }
  txt <- stringr::str_remove_all(txt, pattern = "(?<=pak::pak).*")

  pkgs <- stringr::str_extract_all(
    string = txt,
    pattern = "((?<=(library|require)\\()[a-zA-Z0-9\\.]{1,}(?=(\\)|,))|[a-zA-Z0-9\\.]{1,}(?=\\:{2}))",
    simplify = TRUE
  )

  unique(pkgs[nzchar(pkgs)])
}

pkgs <- tibble(
  script = list.files(pattern = ".R$", recursive = TRUE, full.names = TRUE),
  Lesson = dirname(script) |> str_remove("^\\./"),
  Packages = map(script, script_deps, comment = TRUE)
)

pkgs_info <- unlist(pkgs$Packages) |>
  unique() |>
  setdiff(tidyverse::tidyverse_packages()) |>
  setdiff(tidymodels::tidymodels_packages()) |>
  setdiff(easystats::easystats_packages()) |>
  setdiff(rownames(installed.packages(priority = "base"))) |>
  c("tidyverse", "tidymodels", "easystats") |>
  sort() |>
  pak::pkg_status()
```

```{r}
pkgs |>
  summarize(
    Packages = list(unique(unlist(Packages))),
    .by = Lesson
  ) |>
  reframe(
    Lesson = glue("[{Lesson}]({Lesson}//)"),
    Packages = sapply(Packages, \(x) glue('`{x}`') |> glue_collapse(sep = ", "))
  ) |>
  knitr::kable()
```

<details><summary><i>Installing R Packages</i></summary>

You can install all the R packages used by running:

```{r}
cat("# in alphabetical order:\n\n")

cat("pak::pak(\n  c(")

pkgs_info |>
  transmute(
    prefix = case_when(
      repotype %in% c("cran", "bioc") ~ glue("{package}"),
      remotetype == "github" ~ glue("github::{remotepkgref}")
    ),
    comma = rep(c(",", ""), c(nrow(pkgs_info) - 1, 1)),
    version = glue("# {version}")
  ) |>
  glue_data('    "{prefix}"{comma} {version}') |>
  print()

cat("  )\n)")
```

</details>

---

### Additional Versions

- Prior to 2025, this course was based on the `{caret}` package - this version can still be found [here](https://github.com/mattansb/Machine-Learning-foR-Psychologists/tree/caret).

- Partial parallel `python` lessons can be found in the [`py` folder](py/).

---

### Acknowledgements

Materials developed with Yael Bar-Shachar.
